Universals of word order result from optimization of grammars for efficient communication

### 2019-01-23 writeup/paper-names.tex
unified color scheme (which may require reworking)

@@

\begin{dependency}[theme=simple]
    \begin{deptext}
    I \& wrote \& a \& letter \& to \& a \& friend \\
    \end{deptext}
    \depedge{2}{1}{nsubj}
    \depedge{2}{4}{obj}
    \depedge{4}{3}{ }
    \depedge{4}{3}{ }
    \depedge{2}{5}{}
    \depedge{5}{7}{ }
    \depedge{7}{6}{ }
\end{dependency}
	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right). We provide cross-entropy for predictability (lower is better) and unlabeled attachment scores---a standard metric from natural language processing---for parseability (higher is better).}\label{fig:predict-pars}
\end{figure*}

Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency. Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency. We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}. In 3/4 of the cases shown, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in Chinese and Hindi, the real language is best matched by efficiency optimization.

@@

\subsection{Greenbergian Word Order Correlations}
\label{sec:greenberg}
\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner' (the `Correlate' column), whose relative order correlates with that of verbs and objects.
	For clarity, we organize the correlations by the category of the `verb patterner'.
	For each correlation, we give our operationalization in terms of UD. A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation. 
	For each correlation we report how many of the languages in our sample satisfied it (`Real').
We then report, for each correlation and each objective function, how many of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
\end{table*}

@@

We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages. To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks. The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) in all except four languages, for WALS did not have data.

@@

	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
    \label{fig:posterior-satisfied-count}
\end{figure}

@@

\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, and audiences at CAMP 2018 for helpful discussion.}

>>>>

% TODO where do we say that we're basing our work on data from 51 languages?

@@

%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    I \& wrote \& a \& letter \& to \& a \& friend \\
%    \end{deptext}
%    \depedge{2}{1}{nsubj}
%    \depedge{2}{4}{obj}
%    \depedge{4}{3}{ }
%    \depedge{4}{3}{ }
%    \depedge{2}{5}{}
%    \depedge{5}{7}{ }
%    \depedge{7}{6}{ }
%\end{dependency}
	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right). We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better). Both measures are normalized by sentence length.}\label{fig:predict-pars}
\end{figure*}

Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency. Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency. We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}. In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.

@@

\subsection{Greenbergian Word Order Correlations}
\label{sec:greenberg}
\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
%	For clarity, we organize the correlations by the category of the `verb patterner'.
	For each correlation, we give our operationalization in terms of UD. %A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation. 
	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
\end{table*}

@@

We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages. To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks. The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS. %in all except four languages, for WALS did not have data.

@@

	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies. Efficiency predicts all ten correlations with high posterior probability.}
    \label{fig:posterior-satisfied-count}
\end{figure}

@@

\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
